apply flashr with OCV

### 2018-02-11 analysis/simulation3.Rmd
Files commited by wflow_commit.

---
title: "simulation for the sparse factor loadings"
author: "Wei Wang"
date: 2017-04-04
output: 
  html_document:
    code_folding: hide
---

@@

flashBF.wrapper = function(Y_data,K = 3){
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  library(ebnm)
  library(flashr2)
  data =  flashr2::flash_set_data(Y)
  f_greedy = flashr2::flash_add_greedy(data,verbose=F,var_type = "constant", K = K)
  f = flashr2::flash_backfit(data,f_greedy,var_type = "constant")
  Y_hat = f$EL %*% t(f$EF)
  return(Y_hat)
}

flash.wrapper = function(Y_data, K = 3){
  # missindex is a matirx with 3 column here: i j x
  # Y has miss value already
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  data =  flashr2::flash_set_data(Y)
  g_flash = flashr2::flash_add_greedy(data,verbose=F,var_type = "constant", K = K)
  Y_hat = g_flash$EL %*% t(g_flash$EF)
  return(Y_hat)
}
```

```{r}
library(PMA)
library(flashr2)
library(ssvd)
library(softImpute)
Y_data = DataMaker(seed_num = 0)

LF_1 = PMA.wrapper(Y_data)
LF_2 = flash.wrapper(Y_data)
LF_3 = flashBF.wrapper(Y_data)
LF_4 = SVD.wrapper(Y_data)
LF_5 = SSVD.wrapper(Y_data)
LF_6 = softImpute.wrapper(Y_data)
LF_true = Y_data$L_true %*% t(Y_data$F_true)

library(reshape2)
melted_cormat_1 <- melt(LF_1, na.rm = TRUE)
melted_cormat_2 <- melt(LF_2, na.rm = TRUE)
melted_cormat_3 <- melt(LF_3, na.rm = TRUE)
melted_cormat_4 <- melt(LF_4, na.rm = TRUE)
melted_cormat_5 <- melt(LF_5, na.rm = TRUE)
melted_cormat_6 <- melt(LF_6, na.rm = TRUE)
melted_cormat_true <- melt(LF_true, na.rm = TRUE)
# Heatmap
library(ggplot2)
plot_HP = function(title_name = "truth", melted_cormat){
  p1 = ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
   geom_tile(color = "white")+
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
     midpoint = 0, limit = c(-2,2), space = "Lab") + labs(title = title_name, y = "samples", x = "variables") +
    theme_minimal() + theme(legend.position="none") 
  p1
}

p0 = plot_HP(title_name = "Truth", melted_cormat = melted_cormat_true)
p1 = plot_HP(title_name = "PMD.cv1", melted_cormat = melted_cormat_1)
p2 = plot_HP(title_name = "flash", melted_cormat = melted_cormat_2)
p3 = plot_HP(title_name = "flashb", melted_cormat = melted_cormat_3)
p4 = plot_HP(title_name = "SVD", melted_cormat = melted_cormat_4)
p5 = plot_HP(title_name = "SSVD", melted_cormat = melted_cormat_5)
p6 = plot_HP(title_name = "SI.cv", melted_cormat = melted_cormat_6)
```

```{r}
gridExtra::grid.arrange(p0,p5,p6,p4,p1,p2,ncol = 2)
```

>>>>

---
title: "simulation for the sparse factor loadings"
author: "Wei Wang"
output: 
  html_document:
    code_folding: hide
---

@@

flashBF.wrapper = function(Y_data,K = 3){
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  library(ebnm)
  library(flashr)
  data =  flashr::flash_set_data(Y)
  f_greedy = flashr::flash_add_greedy(data,verbose=F,var_type = "constant", K = K)
  f = flashr::flash_backfit(data,f_greedy,var_type = "constant")
  Y_hat = f$EL %*% t(f$EF)
  return(Y_hat)
}

flash.wrapper = function(Y_data, K = 3){
  # missindex is a matirx with 3 column here: i j x
  # Y has miss value already
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  data =  flashr::flash_set_data(Y)
  g_flash = flashr::flash_add_greedy(data,verbose=F,var_type = "constant", K = K)
  Y_hat = g_flash$EL %*% t(g_flash$EF)
  return(Y_hat)
}
```

## average of 100 simulations

```{r}
HMatrix = readRDS("../data/simulation/rankthree/HeatMap.rds")
T = 100
AVG_HM = matrix(0,nrow = 36000,ncol = 7)
for(i in 1:T){
  AVG_HM = AVG_HM + abs(HMatrix[i,,])
}
AVG_HM = AVG_HM/100
```

```{r}
LF_1 = matrix(AVG_HM[,1],ncol = 240, nrow = 150)
LF_2 = matrix(AVG_HM[,2],ncol = 240, nrow = 150)
LF_3 = matrix(AVG_HM[,3],ncol = 240, nrow = 150)
LF_4 = matrix(AVG_HM[,4],ncol = 240, nrow = 150)
LF_5 = matrix(AVG_HM[,5],ncol = 240, nrow = 150)
LF_6 = matrix(AVG_HM[,6],ncol = 240, nrow = 150)
LF_true = matrix(AVG_HM[,7],ncol = 240, nrow = 150)
library(reshape2)
melted_cormat_1 <- melt(LF_1, na.rm = TRUE)
melted_cormat_2 <- melt(LF_2, na.rm = TRUE)
melted_cormat_3 <- melt(LF_3, na.rm = TRUE)
melted_cormat_4 <- melt(LF_4, na.rm = TRUE)
melted_cormat_5 <- melt(LF_5, na.rm = TRUE)
melted_cormat_6 <- melt(LF_6, na.rm = TRUE)
melted_cormat_true <- melt(LF_true, na.rm = TRUE)
# Heatmap
library(ggplot2)
plot_HP = function(title_name = "truth", melted_cormat){
  p1 = ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
   geom_tile(color = "white")+
    #scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
     #midpoint = 0, limit = c(-0.81,0.81), space = "Lab") + labs(title = title_name, y = "samples", x = "variables") +
    scale_fill_gradient2(low = "white", high = "grey6", mid = "grey9", 
     midpoint = 0.41, limit = c(0,0.82), space = "Lab") + labs(title = title_name, y = "samples", x = "variables") +
    theme_minimal() + theme(legend.position="none",plot.title = element_text(size = 12.9, face = "bold")) 
  p1
}

p0 = plot_HP(title_name = "Latent structure (truth)", melted_cormat = melted_cormat_true)
p1 = plot_HP(title_name = "PMD.cv1", melted_cormat = melted_cormat_1)
p2 = plot_HP(title_name = "flash", melted_cormat = melted_cormat_2)
p3 = plot_HP(title_name = "flash_pn", melted_cormat = melted_cormat_3)
p4 = plot_HP(title_name = "SVD", melted_cormat = melted_cormat_4)
p5 = plot_HP(title_name = "SSVD", melted_cormat = melted_cormat_5)
p6 = plot_HP(title_name = "SI.cv", melted_cormat = melted_cormat_6)

```

```{r}
gridExtra::grid.arrange(p2,p3,p1,p6,p5,p4,ncol = 3)
```

### 2018-02-11 analysis/templateforOCV.Rmd
Files commited by wflow_commit.

---
title: "apply flashr2 with OCV"
author: "wei wang"
date: 09-30
output: 
  html_document:
    code_folding: hide
---

@@

call_method = function(Y_data,method,Kmax = 50){
  if(method == "flash"){
    data = flashr2::flash_set_data(Y_data)
    f_greedy = flashr2::flash_add_greedy(data,K=Kmax)
    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
  }else if(method == "flash_wn"){
    data = flashr2::flash_set_data(Y_data)
    f_greedy = flashr2::flash_add_greedy(data,K=Kmax,nullcheck=FALSE)
    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
  }else if(method == "flash_gb"){
    data = flashr2::flash_set_data(Y_data)
    f_greedy = flashr2::flash_add_greedy(data,K=Kmax)
    f = flashr2::flash_backfit(data,f_greedy)
    Y_hat = f$EL %*% t(f$EF)
  }else if(method == "flash_bf"){
    data = flashr2::flash_set_data(Y_data)
    f_data = flashr2::flash_add_factors_from_data(data,K = Kmax)
    f = flashr2::flash_backfit(data,f_data)
    Y_hat = f$EL %*% t(f$EF)
  }else if(method == "PMD"){
    res_log = capture.output({out = PMA.wrapper(Y_data,ngrids = 10,K = Kmax,fold = 10)})
    if(length(out$d)==1){
      Y_hat =  (out$d) * out$u %*% t(out$v)
    }else{
      Y_hat =  out$u %*% diag(out$d) %*% t(out$v)
    }
  }else if(method == 'PN'){
    library(ebnm)
    library(flashr2)
    data = flashr2::flash_set_data(Y_data)
    f_greedy = flashr2::flash_add_greedy(data,K=Kmax,ebnm_fn = ebnm_pn)
    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
  }else if(method == 'PMD_d'){
    out = PMA::PMD(Y_data, K = Kmax)
    Y_hat =  out$u %*% diag(out$d) %*% t(out$v)
  }else if(method == 'softImpute_d'){
    gsoft = softImpute::softImpute(Y_data, rank.max = Kmax)
    Y_hat =  gsoft$u %*% diag(gsoft$d) %*% t(gsoft$v)
  }else if(method == "softImpute"){
    gsoft = try(softImpute.wrapper(Y_data,ngrids = 10,K = Kmax,fold = 10))
    if(length(gsoft$d)==1){
      Y_hat =  try((gsoft$d) * gsoft$u %*% t(gsoft$v))
    }else{
      Y_hat =  try(gsoft$u %*% diag(gsoft$d) %*% t(gsoft$v))
    }
  }else if(method == "SF_flash"){
    gsoft = try(softImpute.wrapper(Y_data,ngrids = 10,K = Kmax,fold = 10))
    LL = gsoft$u %*% diag(gsoft$d)
    FF = gsoft$v
    data = flashr2::flash_set_data(Y_data)
    f_data = flashr2::flash_add_lf(data, LL=LL, FF=FF)
    f = flashr2::flash_backfit(data,f_data)
    Y_hat = f$EL %*% t(f$EF)
  }else{
    # stop("the method is not implemented yet, please check it out")
    Y_hat = matrix(0,ncol = dim(Y_data)[2],nrow = dim(Y_data)[1])
  }
  return(Y_hat)
}

@@

In this case, we have problem in running `flashr2` on RCC. so we just try other methods first.

@@

result = c(PN_rmse,flashG_rmse,flashGwn_rmse,flashB_rmse,nsfa_rmse,pmd_rmse,soft_rmse)
saveRDS(result, "./NBSFAout/output.rds")
```

@@

library(ggplot2)
plot_res = function(output,title = "data",legend_position = "none", x_label){
  rmse = as.vector(output)
  N = dim(output)[1]
  methods = rep(x_label, each = N)
  df = data.frame(RMSE = rmse, Method = methods )
  p<-ggplot(df, aes(x=Method, y=RMSE, color=Method)) +
    geom_boxplot()+
    # geom_violin()+
    ggtitle(title) +  theme_bw()+
    theme(legend.position= legend_position, legend.text=element_text(size=15),
          plot.title = element_text(size = 15, face = "bold"),
          axis.text.y = element_text(size =12.6),
          axis.text.x = element_text(size =12.6,angle = 45, hjust = 1))
  p
}

@@

@@

x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotM = plot_res(output,"MovieLens data",x_label = x_label)

@@

# the result order would be flashG PMD PMDcv NSFA SF SFcv
output = cbind(bc_res[,2],bc_res[,1],BC_res[,3],bc_res[,6],BC_res[,2],BC_res[,4],bc_res[,7])
x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotB = plot_res(output,"Breast Cancer data",x_label = x_label)

@@

# the result order would be flashG PMD PMDcv NSFA SF SFcv
output = cbind(gz_res[,1],gz_pn_res[,2],GZ_res[,3],gz_res[,3],GZ_res[,2],GZ_res[,4],gz_res[,4])
x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotG = plot_res(output,"GTEx data",x_label = x_label)

@@

# the result order would be flashG PMD PMDcv NSFA SF SFcv
output = cbind(pt_res[,3],pt_res[,1],PT_res[,3],pt_res[,6],PT_res[,2],PT_res[,4],pt_res[,7])
x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotP = plot_res(output,"Text data",x_label = x_label)

@@

# the result order would be flashG PMD PMDcv NSFA SF SFcv
output = cbind(dt_res[,2],dt_res[,1],DT_res[,3],dt_res[,6],DT_res[,2],DT_res[,4],dt_res[,7])
x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotT = plot_res(output,"Tumor data",x_label = x_label)

# gridExtra::grid.arrange(plotP,plotT,plotG,plotB,plotM, layout_matrix = rbind(c(1,NA,2),c(NA,5,NA),c(4,NA,3)))
gridExtra::grid.arrange(plotM,plotT,plotG,plotB,plotP, layout_matrix = rbind(c(1,2,3),c(4,5,NA) ))

>>>>

---
title: "apply flashr with OCV"
author: "wei wang"
output: 
  html_document:
    code_folding: hide
---

@@

call_method = function(Y_data,method,Kmax = 50){
  if(method == "flash"){
    data = flashr::flash_set_data(Y_data)
    f_greedy = flashr::flash_add_greedy(data,K=Kmax)
    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
  }else if(method == "flash_wn"){
    data = flashr::flash_set_data(Y_data)
    f_greedy = flashr::flash_add_greedy(data,K=Kmax,nullcheck=FALSE)
    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
  }else if(method == "flash_gb"){
    data = flashr::flash_set_data(Y_data)
    f_greedy = flashr::flash_add_greedy(data,K=Kmax)
    f = flashr::flash_backfit(data,f_greedy)
    Y_hat = f$EL %*% t(f$EF)
  }else if(method == "flash_bf"){
    data = flashr::flash_set_data(Y_data)
    f_data = flashr::flash_add_factors_from_data(data,K = Kmax)
    f = flashr::flash_backfit(data,f_data)
    Y_hat = f$EL %*% t(f$EF)
  }else if(method == "PMD"){
    res_log = capture.output({out = PMA.wrapper(Y_data,ngrids = 10,K = Kmax,fold = 10)})
    if(length(out$d)==1){
      Y_hat =  (out$d) * out$u %*% t(out$v)
    }else{
      Y_hat =  out$u %*% diag(out$d) %*% t(out$v)
    }
  }else if(method == 'PN'){
    library(ebnm)
    library(flashr)
    data = flashr::flash_set_data(Y_data)
    f_greedy = flashr::flash_add_greedy(data,K=Kmax,ebnm_fn = ebnm_pn)
    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
  }else if(method == 'PMD_d'){
    out = PMA::PMD(Y_data, K = Kmax)
    Y_hat =  out$u %*% diag(out$d) %*% t(out$v)
  }else if(method == 'softImpute_d'){
    gsoft = softImpute::softImpute(Y_data, rank.max = Kmax)
    Y_hat =  gsoft$u %*% diag(gsoft$d) %*% t(gsoft$v)
  }else if(method == "softImpute"){
    gsoft = try(softImpute.wrapper(Y_data,ngrids = 10,K = Kmax,fold = 10))
    if(length(gsoft$d)==1){
      Y_hat =  try((gsoft$d) * gsoft$u %*% t(gsoft$v))
    }else{
      Y_hat =  try(gsoft$u %*% diag(gsoft$d) %*% t(gsoft$v))
    }
  }else if(method == "SF_flash"){
    gsoft = try(softImpute.wrapper(Y_data,ngrids = 10,K = Kmax,fold = 10))
    LL = gsoft$u %*% diag(gsoft$d)
    FF = gsoft$v
    data = flashr::flash_set_data(Y_data)
    f_data = flashr::flash_add_lf(data, LL=LL, FF=FF)
    f = flashr::flash_backfit(data,f_data)
    Y_hat = f$EL %*% t(f$EF)
  }else{
    # stop("the method is not implemented yet, please check it out")
    Y_hat = matrix(0,ncol = dim(Y_data)[2],nrow = dim(Y_data)[1])
  }
  return(Y_hat)
}

@@

In this case, we have problem in running `flashr` on RCC. so we just try other methods first.

@@

result = c(PN_rmse,flashG_rmse,flashGwn_rmse,flashB_rmse,nsfa_rmse,pmd_rmse,soft_rmse)
saveRDS(result, "./NBSFAout/output.rds")
print(sessionInfo())
```

@@

library(ggplot2)
plot_res = function(output,title = "data",legend_position = "none", x_label, myColors){
  rmse = as.vector(output)
  N = dim(output)[1]
  methods = rep(x_label, each = N)
  df = data.frame(RMSE = rmse, Method = methods )
  p<-ggplot(df, aes(x=Method, y=RMSE, color=Method)) +
    geom_boxplot()+
    # geom_violin()+
    ggtitle(title) +  theme_bw()+  scale_color_manual(values=myColors)+
    theme(legend.position= legend_position, legend.text=element_text(size=15),
          plot.title = element_text(size = 15, face = "bold"),
          axis.text.y = element_text(size =12.6),
          axis.text.x = element_text(size =12.6,angle = 45, hjust = 1))
  p
}

@@

fill_color = c("red","yellow3", "tan1","springgreen3", "springgreen","deepskyblue3","deepskyblue")

@@

x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotM = plot_res(output,"MovieLens data",x_label = x_label,myColors = fill_color)

@@

# the result order would be flashG PMD PMDcv NSFA SF SFcv
output = cbind(bc_res[,2],bc_res[,1],BC_res[,3],bc_res[,6],BC_res[,2],BC_res[,4],bc_res[,7])
x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotB = plot_res(output,"Breast cancer data",x_label = x_label,myColors = fill_color)

@@

# the result order would be flashG PMD PMDcv NSFA SF SFcv
output = cbind(gz_res[,1],gz_pn_res[,2],GZ_res[,3],gz_res[,3],GZ_res[,2],GZ_res[,4],gz_res[,4])
x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotG = plot_res(output,"GTEx data",x_label = x_label,myColors = fill_color)

@@

# the result order would be flashG PMD PMDcv NSFA SF SFcv
output = cbind(pt_res[,3],pt_res[,1],PT_res[,3],pt_res[,6],PT_res[,2],PT_res[,4],pt_res[,7])
x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotP = plot_res(output,"Presidential address data",x_label = x_label,myColors = fill_color)

@@

# the result order would be flashG PMD PMDcv NSFA SF SFcv
output = cbind(dt_res[,2],dt_res[,1],DT_res[,3],dt_res[,6],DT_res[,2],DT_res[,4],dt_res[,7])
x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotT = plot_res(output,"Tumor data",x_label = x_label,myColors = fill_color)
```

```{r}
# gridExtra::grid.arrange(plotP,plotT,plotG,plotB,plotM, layout_matrix = rbind(c(1,NA,2),c(NA,5,NA),c(4,NA,3)))
gridExtra::grid.arrange(plotM,plotT,plotG,plotB,plotP, layout_matrix = rbind(c(1,2,3),c(4,5,NA) ))
```
